Constant-round Blind Classical Verification of Quantum Sampling

### 2020-04-15 qpip0.tex
Add action items on 8.2

In her groundbreaking work, Mahadev~\cite{FOCS:Mahadev18a} gave a 4-round $\QPIP_0$ protocol to let a $\BQP$ machine "commit a $XZ$ measurement" to a classical machine.
Intuitively, the protocol allows the prover to commit to a state $\rho$ and the verifier to choose a measurement basis $h$ and learn a corresponding measurement outcome $M_{XZ}(\rho, h)$.
\Ethan{The following discussion is from Han-Hsuan}
Note that even a honest prover has complete freedom in choosing $\rho$;
what this protocol actually does is let the prover commit to a particular measurement on $\rho$.
This is good enough for our application, since we will feed the $XZ$ measurement results into the verification of $\QPIP_1$ protocol to ensure that the prover pick the history state as $\rho$ as we will show next.
We now describe the protocol while omitting the implementation details.

@@

Protocol:
\begin{enumerate}
	\item The verifier generates a public and secret key pair $(pk, sk)\leftarrow\cVMeasure{1}(1^\lambda, h)$. It sends $pk$ to the prover.
	\item The prover generates $(y, \sigma)\leftarrow\cPMeasure{2}(pk, \rho)$.
		$y$ is a classical ``commitment", and $\sigma$ is some internal state.
		He sends $y$ to the verifier.
	\item The verifier samples $c\xleftarrow{\$}\zo$ uniformly at random and sends it to the prover. $c=0$ indicates what's called a ``testing round", and $c=1$ indicates a ``Hadamard round".
	\item The prover generates a classical string $a\leftarrow\cPMeasure{4}(c, \sigma)$ and sends it to back to the verifier.
	\item If it's a testing round ($c=0$), then the verifier generates and outputs $o\leftarrow\cVMeasure{T}(pk, y, a)$, where $o\in\set{\Acc, \Rej}$.
		If it's a Hadamard round ($c=1$), then the verifier computes $M_{XZ}(\rho, h)$ by $v\leftarrow\cVMeasure{H}(sk, h, y, a)$. It outputs $v$.
\end{enumerate}
\end{protocol}

Notice that the verifier doesn't in fact receive a measurement result on a testing round.
Instead, it exists to achieve a ``binding" property that gives certain guarantees against cheating provers.

@@

\Ethan{Again change all n to lambda for sec. param. and fix variable labels}

\begin{lemma}\label{lem:urmila-binding}[binding property of Protocol~\ref{proto:urmila4}\cite{FOCS:Mahadev18a}]
If a prover's strategy in $\PiMeasure$ \hannote{need clarification} results in passing the testing round with probability $1-\negl(n)$, then there must exist a $\rho$ such that for all $h$, the verifier will produce a $z$ on the Hadamard round that is $\negl(n)$-computationally indistinguishable\footnote{Recall that quantum computational indistinguishability is defined in Definition~\ref{def:comp-indis}.} from $M_{XZ}(\rho,h)$, unless LWE can be efficiently solved by a quantum computer. 
\end{lemma}

We also mention a fact that intuitively captures the completeness of $\PiMeasure$.

\begin{fact}[\cite{FOCS:Mahadev18a}]~\label{lem:trivial-4-round-strategy}
   There exist a prover strategy for $\PiMeasure$ that is accepted with probability $1-\negl{\lambda}$ \Ethan{Changed n to lambda here, see what does it affect} in the testing.
\end{fact}

@@

We now use $\PiMeasure$ to modify our $\QPIP_1$ Protocol for $\SampBQP$, $\PiSamp=(\PSamp, \VSamp)$, to achieve a corresponding $\QPIP_0$ protocol $\PiNaive$.
Recall that $\PiSamp$ is a one-message protocol, where $\VSamp$ applies $X$ and $Z$ measurements non-adaptively to the incoming message.
We modify $\VSamp$ to pick all its measurement ahead of time, and measure the incoming message from $\PSamp$ through $\PiMeasure$ instead as below.

@@

	Protocol:
	\begin{enumerate}
		\item The verifier choose a $XZ$-measurement $h\leftarrow\cVNaive{1}(1^\lambda, 1^\varepsilon, x)$ from the distribution specified in Step~\ref{step:qpip1-verify} of $\PiSamp$.
		It picks between testing and Hadamard round by sampling $c\xleftarrow{\$}\zo$ uniformly at random.
	\item The prover prepares $\sigma\leftarrow\cPNaive{2}(1^\lambda, 1^\varepsilon, x)$ by running Step~\ref{step:qpip1-state-gen} of $\PiSamp$.
		\item \label{step:urmila-in-naive} The verifier and prover run $(\PMeasure(\sigma), \VMeasure(h))(1^\lambda, 1^{1/\varepsilon}, x)$.
	\item If it's a testing round, the verifier accepts or rejects based on the outputs of $\PiMeasure$.
		If it's a Hadamard round, the verifier obtains $v$. It finishes the verification step of Protocol~\ref{ProtoQPIP1} by using $v$ as the measurement result he would get and outputs $(d, z)$.
	\end{enumerate}
\end{protocol}

$\PiNaive$ has no soundness guarantee, but it has the following binding property that would be useful in proving the soundness of the actual protocol in Section~\ref{sec:qpip0}.

\begin{lemma}\label{lem:naive-qpip0-binding}[binding property of  Protocol~\ref{proto:qpip0_naive} ]
	\Ethan{I haven't done anything with this lemma yet. Might need to fix or make consistent later.}
If a prover's strategy in Protocol~\ref{proto:qpip0_naive} results in passing the testing round with probability $1-\negl(n)$, then conditioning on the Hadamard round being chosen,

$$\norm{(d,z_{naive})-(d,z_{ideal})}_q \leq O(\eps),$$

unless LWE can be efficiently solved by a quantum computer. 
\end{lemma}

\begin{proof}
By Lemma~\ref{lem:urmila-binding}, if the prover passes the testing round with probability $1-\negl(n)$, there exist a $\rho$ such that $\norm{z_{commit}- M_{XZ}(\rho,h)}_q \leq \negl(n)$. Note that the prover can choose to abort, so $z_{commit}$ and $\rho$ are subnormalized. Let $Ver$ be the verification operation run by Protocol~\ref{ProtoQPIP1} on the measurement outcome. \hannote{describe better} Since $Ver$ is an efficient process and $Ver(z_{commmit})+p_{abort}(\Rej,\bot)=(d,z_{naive})$, we have that

@@

>>>>

\Ethan{Half of this discussion is from Han-Hsuan}

In her groundbreaking work, Mahadev~\cite{FOCS:Mahadev18a} gave a 4-round $\QPIP_0$ protocol to let a $\BQP$ machine ``commit a $XZ$ measurement" to a classical machine.
\Ethan{Mahadev's end result is QPIP0 BQP.
Two steps. One is measurement proto.
The other is combine with MF.
Break the previous sentence into two sentences accordingly. Something like the following:

Mahadev's end results in groundbreaking work 4 round protocol for BQP. At the core it is a 4 round protocol to let BQP prover commit XZ measurement to a classical machine.
}
Intuitively, the protocol allows the prover to commit to a state $\rho$ and the verifier to choose a measurement basis $h$ and learn a corresponding measurement outcome $M_{XZ}(\rho, h)$ \Ethan{Actually define this notation by cutting-pasting $h$ from verifier's input}.
\Ethan{Remove the next sentences or move to later}
Note that even a honest prover has complete freedom in choosing $\rho$;
what this protocol actually does is let the prover commit to a particular measurement on $\rho$.
This is good enough for our application, since we will feed the measurement results into the verification of $\QPIP_1$ protocol to ensure that the prover pick the history state as $\rho$ as we will show next.
We now describe the protocol while omitting the implementation details.

@@

Protocol:
\begin{enumerate}
	\item The verifier generates a public and secret key pair $(pk, sk)\leftarrow\cVMeasure{1}(1^\lambda, h)$. It sends $pk$ to the prover.
	\item The prover generates $(y, \sigma)\leftarrow\cPMeasure{2}(pk, \rho)$.
		$y$ is a classical ``commitment", and $\sigma$ is some internal state.
		He sends $y$ to the verifier.
	\item The verifier samples $c\xleftarrow{\$}\zo$ uniformly at random and sends it to the prover. $c=0$ indicates what's called a ``testing round", and $c=1$ indicates a ``Hadamard round".
	\item The prover generates a classical string $a\leftarrow\cPMeasure{4}(pk, c, \sigma)$ and sends it to back to the verifier.
	\item If it's a testing round ($c=0$), then the verifier generates and outputs $o\leftarrow\cVMeasure{T}(pk, y, a)$, where $o\in\set{\Acc, \Rej}$. \Ethan{Just say testing and acc/rej}
		If it's \Ethan{No abbreviations} a Hadamard round ($c=1$), then the verifier generates and outputs $v\leftarrow\cVMeasure{H}(sk, h, y, a)$.
\end{enumerate}
\end{protocol}

The verifier only learns the measurement outcome on a Hadamard round.
The protocol achieves a ``binding" property that gives guarantees against cheating provers.

@@

\begin{lemma}\label{lem:urmila-binding}[binding property of Protocol~\ref{proto:urmila4}\cite{FOCS:Mahadev18a}]
\Ethan{Need to be more formal. BQP adversaries... etc. Look at the two parallel rep papers}
\Ethan{Changed $n$ to $\lambda$, TODO sync later uses of this lemma}
If a prover's strategy in $\PiMeasure$ \hannote{need clarification} results in passing the testing round with probability $1-\negl(\lambda)$, then there must exist a $\rho$ such that for all $h$, the verifier will produce a $v$ on the Hadamard round that is $\negl(\lambda)$-computationally indistinguishable\footnote{Recall that quantum computational indistinguishability is defined in Definition~\ref{def:comp-indis}.} from $M_{XZ}(\rho,h)$, unless LWE can be efficiently solved by a quantum computer. 
\end{lemma}

We also mention a fact that will be useful later.
\begin{fact}
\label{lem:trivial-4-round-strategy}
   There exist a prover strategy for $\PiMeasure$ that is accepted with probability $1-\negl{\lambda}$ \Ethan{Changed n to lambda here, see what does it affect} in the testing.
\end{fact}

@@

We now use $\PiMeasure$ to transform our $\QPIP_1$ Protocol for $\SampBQP$, $\PiSamp=(\PSamp, \VSamp)$, to a corresponding $\QPIP_0$ protocol $\PiNaive$ in a natural way.
Recall that in $\PiSamp$ the verifier learns the XZ measurement outcome of the prover's message.
In PiInt We let the verifier use $\PiMeasure$ to learn...

@@

	Protocol:
	\begin{enumerate}
		\item The verifier choose a $XZ$-measurement $h\leftarrow\cVNaive{1}(1^\lambda, 1^{1/\varepsilon}, x)$ \Ethan{Just don't give names for now} from the distribution specified in Step~\ref{step:qpip1-verify} of $\PiSamp$.
	\item The prover prepares $\rho\leftarrow\cPNaive{2}(1^\lambda, 1^{1/\varepsilon}, x)$ by running Step~\ref{step:qpip1-state-gen} of $\PiSamp$.
		\item \label{step:urmila-in-naive} The verifier and prover run $(\PMeasure(\rho), \VMeasure(h))(1^\lambda)$.
		If it's a testing round, the verifier accepts or rejects based on the outcome of $\PiMeasure$.
		If it's a Hadamard round, the verifier obtains $v$.
	\item 
		If it's a Hadamard round, the verifier finishes the verification step of Protocol~\ref{ProtoQPIP1} by generating and outputting $(d, z)\leftarrow\cVNaive{4}(1^\lambda, 1^{1/\varepsilon}, x, h, v)$.
	\end{enumerate}
\end{protocol}

$\PiNaive$ has no soundness guarantee \Ethan{Sounds strange}, but it has the following binding property that would be useful in proving the soundness of the actual protocol in Section~\ref{sec:qpip0}.

\Ethan{
Note that in PiInt the verifier only learns the output in H round
So PiInt is not QPIP0 with good completeness or soundness.
On the other hand, it has the following binding property (maybe refer that it does because PiMeasure does)
}

\begin{lemma}\label{lem:naive-qpip0-binding}[binding property of $\PiNaive$]
\Ethan{Statement needs to be more rigorous}
If a prover's strategy in Protocol~\ref{proto:qpip0_naive} results in passing the testing round with probability $1-\negl(n)$, then conditioning on the Hadamard round being chosen,

\Ethan{Instead of using conditioning, just say the verifier's output in the Hadamard round satisfies this}

$$(d, z)\approx_{c, \varepsilon}(d, z_{ideal})$$

unless LWE can be efficiently solved by a quantum computer.
\Ethan{under the QLWE assumption}
\end{lemma}

\begin{proof}
\Ethan{I haven't looked at this yet}
By Lemma~\ref{lem:urmila-binding}, if the prover passes the testing round with probability $1-\negl(n)$, there exist a $\rho$ such that $\norm{z_{commit}- M_{XZ}(\rho,h)}_q \leq \negl(n)$. Note that the prover can choose to abort, so $z_{commit}$ and $\rho$ are subnormalized. Let $Ver$ be the verification operation run by Protocol~\ref{ProtoQPIP1} on the measurement outcome. \hannote{describe better} Since $Ver$ is an efficient process and $Ver(z_{commmit})+p_{abort}(\Rej,\bot)=(d,z_{naive})$, we have that

@@

\Ethan{Define in words first, and use the protocol box to introduce notations

unfold for the sake of convenience in analysis

verifier h1 to hm, prover generates whatever it should generate (tensor of whatever),
pki ski expand all, use vec.

}
